Symbolic Weighted Parsing and\\ Automated Music Transcription

### 2021-05-03 main.tex
intro

% Parsing over infinite alphabet as optimal alignment computation 
% as edit-distance between string and language
% 
\title{Symbolic Parsing and\\ Automated Music Transcription}
%\title{Weighted Visibly Pushdown Automata\\ Automated Music Transcription}
\author{Florent Jacquemard}
%\institute{INRIA \& CNAM, Paris, France\\
%\email{florent.jacquemard@inria.fr}}

@@

Weighted parsing~\cite{Goodman99SemiringParsing,Nederhof03weightedParsing,MorbitzVogler19weighted-parsing} 
consists in analyzing a word in input \wrt a weighted language model, \eg a probabilistic context-free grammar (CFG).
In such a model, a weight value is associated to each transition rule, % production rule
and the rule's weights can be combined with a associative product operator $\otimes$ into 
the weight of a derivation (or \emph{abstract syntax tree}, \ie the result of an analyze).
A second operator $\oplus$, associative and commutative, 
is used to handle ambiguity of the model, 
by summing the weights of the possibly several (in general exponentially many) 
syntax trees associated to a given input word.
Intuitively, $\oplus$ aims at selecting, or ranking, the syntax trees.
The weight domain, equipped with these two operators is assumed, at minima, 
to form a \emph{semiring} $\Semiring$. %\cite{Eilenberg74automata}
%by ranking 
%making the weight domain a semiring.
Some efficient specialized parsing algorithms~\cite{Huang05kbest} have been proposed in this context 
in order to compute the $n$ best syntax trees of a given input word without having to enumerate them all.
Generally based on dynamic programming, these algorithms rely on 
additional algebraic properties of $\Semiring$.
-- see \eg~\cite{Huang05kbest} for some NLP applications.
%The extraction of $n$ best list is useful

The various approaches to weighted parsing usually assume a finite input alphabet. %models and algorithms 
Considering infinite or large alphabets can however be of practical interest, 
for instance when processing strings in UTF encodings or for vulnerability detection in Web-applications 
(see \eg~\cite{dAntoni21CACM}), for model checking, 
or when dealing with sequences of timestamped events
(like in the approach of~\cite{MCM19} of automated music transcription by parsing).
%generalizes weighted parsing: 
%finding the best derivation for a weighted grammar.

We propose here some models of weighted recognizers 
%model of weighted CFG 
computing on words over infinite input alphabets %sets of terminal symbol
- but with finite sets of states and transitions rules. 
% non-terminals and production rules).
In their transition rules, input symbols appear as variables (parameters)
and the weight associated to the transition rule is a function of these variables.
%
This approach is close to the case of Symbolic Automata (SA)~\cite{VeanesdAntoniJACM}, 
except that the domain for weight values is not restricted to be Boolean, 
like for the guards in the rules of SA, 
but can be an arbitrary commutative semiring (assuming some restrictions).
%
The models defined here are finite automata, transducers and pushdown automata with visibly restriction.
%wrt an edit distance
The latter model of these automata computes on nested words~\cite{AlurMadhusudan09nested}, 
a structured format of words parenthesized with markup symbols, 
which can correspond to a linearization of trees.

The latter corresponds in expressiveness
to a strict restricted model of weighted grammars
and can be used in a general approach for parsing over infinite input alphabets.
\dots
First one general edit-distance is defined by a weighted word 
transducer~\cite{Mohri}
%Symbolic automata transducers are extended models~\cite{VeanesdAntoniJACM}
%dealing with infinite set of input symbols...
value in a semiring...

\section{SW Automata and Transducers}
\label{section:transducer}

@@

\section{SW Visibly Pushdown Automata}
\label{section:SWVPA}
The following model generalizes Symbolic VPA~\cite{DAntonyAlur14SVPDA}
from Boolean semirings to arbitrary semiring weight domains.

@@

\section{Application}
Symbolic Automated Music Transcription
and analysis of music performances

>>>>

% Parsing over infinite alphabet as optimal alignment computation 
% as edit-distance between string and language
% 
\title{Symbolic Weighted Parsing and\\ Automated Music Transcription}
%\title{Weighted Visibly Pushdown Automata\\ Automated Music Transcription}
\author{Florent Jacquemard}
%\institute{INRIA \& CNAM, Paris, France\\
%\email{florent.jacquemard@inria.fr}}

@@

Weighted parsing~\cite{Goodman99SemiringParsing,Nederhof03weightedParsing,MorbitzVogler19weighted-parsing} 
consists in analyzing a word in input \wrt a weighted language model, \eg a probabilistic context-free grammar (CFG).
In such a model, a weight value is associated to each transition rule, % production rule
and the rule's weights can be combined with a associative product operator $\otimes$ into 
the weight of a derivation (or \emph{abstract syntax tree}, \ie the result of an analyze).
A second operator $\oplus$, associative and commutative, 
is used to handle ambiguity of the model, 
by summing the weights of the possibly several (in general exponentially many) 
syntax trees associated to a given input word.
Intuitively, $\oplus$ aims at selecting, or ranking, the syntax trees.
The weight domain, equipped with these two operators is assumed, at minima, 
to form a \emph{semiring} $\Semiring$, %\cite{Eilenberg74automata}
such as a min-plus algebra of domain $\mathbb{R} \cup \{ +\infty\}$, 
where $\oplus$ is min and $\otimes$ is plus.
%by ranking 
%making the weight domain a semiring.
Some efficient specialized parsing algorithms~\cite{Huang05kbest} have been proposed in this context 
in order to compute the $n$ best syntax trees of a given input word without having to enumerate them all.
Generally based on dynamic programming, these algorithms rely on 
additional algebraic properties of~$\Semiring$.
-- see \eg~\cite{Huang05kbest} for some NLP applications.
%The extraction of $n$ best list is useful

The various approaches to weighted parsing usually assume a finite input alphabet. %models and algorithms 
Considering infinite or large alphabets can however be of practical interest, 
for instance in the context of processing strings in UTF16 encodings, 
for vulnerability detection in Web-applications 
(see \eg~\cite{dAntoni21CACM}), 
for the representation of execution traces in model checking, 
or when dealing with sequences of timestamped events
-- we describe briefly at the end of the paper a parsing based approach to 
of automated music transcription, that is the conversion of a symbolic 
representation of a music performance into a structured music score.
%\cite{MCM19} 
%generalizes weighted parsing: 
%finding the best derivation for a weighted grammar.

We present here some weighted language models
%model of weighted CFG 
computing on words over infinite input alphabets %sets of terminal symbol
- but with finite sets of states and transitions rules. 
% non-terminals and production rules).
In their transition rules, input symbols appear as variables %(parameters)
and the weight associated to a transition rule is a function of these variables.
% 
This approach is close to the case of Symbolic Automata (SA)~\cite{dAntoni21CACM}, 
except that the domain for weight values is not restricted to be Boolean, 
like for the guards in the rules of SA, 
but can be an arbitrary commutative semiring (assuming some restrictions).
%
The models defined here are finite automata (\SWA), 
transducers (\SWT) and pushdown automata
with a visibly restriction~\cite{AlurMadhusudan09nested} (\SWVPA).
%wrt an edit distance
The latter model of automata computes on \emph{nested words}~\cite{AlurMadhusudan09nested}, 
a structured form of words parenthesized with markup symbols, 
corresponding to a linearization of trees.
In expressiveness, they correspond to a strict restriction
of weighted CFG. 
%and can be used in a general approach for parsing over infinite input alphabets.
Let $A$ be a \SWVPA, associating $A(w) \in \Semiring$ 
to a given a nested word $w$ (representing a parse tree),
and let a \SWT compute a distance $d$, in $\Semiring$, 
between 2 strings over respectively an infinite input alphabet and the 
same (infinite) alphabet of $A$.
Then, the problem of Symbolic Weighted Parsing is, 
given an input string $s$, to find a nested word $w$ minimizing 
(according to the ranking defined by $\oplus$)
the distance $d(s, w) \otimes A(w)$ between $s$ and $A$, 
as defined in~\cite{Mohri03EDWA}.
%following

% First one general edit-distance is defined by a weighted word 
% transducer~\cite{Mohri}
% %Symbolic automata transducers are extended models~\cite{VeanesdAntoniJACM}
% %dealing with infinite set of input symbols...
% value in a semiring...

In Section~\ref{section:SWA} we introduce \SWA and \SWT, 
then \SWVPA are defined in Section~\ref{section:SWPDA}, 
where a polynomial 1-best algorithm is described that can be use to solve
Symbolic Weighted Parsing.
Finally, \label{sec:transcription}, we present an application 
of this approach to automated music transcription that has been implemented.

\section{SW Automata and Transducers}
\label{section:transducer}
\label{section:SWA}
\label{section:SWT}

@@

\section{SW Visibly Pushdown Automata}
\label{section:SWVPA}\label{sec:SWVPA}
The following model generalizes Symbolic VPA~\cite{DAntonyAlur14SVPDA}
from Boolean semirings to arbitrary semiring weight domains.

@@

\section{Application}
\label{sec:transcription}
Symbolic Automated Music Transcription
and analysis of music performances
